R modeling to prepare code for R shiny app exploring monster encounter information with pathfinder 2e 
```{r}
library(tidyverse)
```

```{r}
sim_results <- read.csv('../output_data/simualtion_results_ac10-30.csv')
sim_results|> head(3)
```
Identify rows or columns with blank or inaccurately copied filter columns and drop them (for now)
```{r}
# Replace all empty strings with NA
sim_results <- sim_results  |> mutate(across(c(ancestry, class_build), na_if, ""))

# sim_results |> drop_na(class_build) |> drop_na(ancestry)

sim_results |> distinct(class_build)
```
```{r}
all_class_options <- sim_results|> distinct(class_build) |> pull()
all_class_options

sim_results |> filter(class_build == all_class_options[1])
```


```{r}
# #make a visual where each class shows total damage 
# class_list <- sim_results |> c(distinct(class_build))
# 
# class_list |> pull(class_build)

### list all class options for chosing 
all_class_options <- sim_results|> distinct(class_build) |> pull()

class_choice = all_class_options[4]

#graph results from only one class ### figure out how to automate this process later 
class_select_sim <- sim_results |> filter(class_build == class_choice)

#visualize the total_dmg output for each class split by ancestry 
class_select_sim |> ggplot(aes(x=target_ac, y=total_dmg, color=ancestry)) +  
  geom_smooth(aes(group=ancestry), method='lm', se=TRUE) +
  labs(title=glue('{class_choice} damage output by level and target ac'), 
       x='target ac', 
       y= 'total damage over 5 round')

```

troublehsoot ancestry plot color issue -> make it a factor variable?
```{r}
class_select_sim$ancestry <- factor(class_select_sim$ancestry)
class_select_sim
```

Open monster stat and add the monster HP distribution to the graph 
```{r}

monster_data<- read.csv('../output_data/bestiary_stats.csv')
monster_data|> head(3)

```

Visualize the distribution of total damage needed for specific hitpoint levels
```{r}
#filter so monster data only goes to the max AC level of simulation dataset 
max_target_ac <- sim_results |> distinct(target_ac) |> max()
  
monster_data_maxac <- monster_data |> filter(AC <= max_target_ac)

monster_data_maxac|> head()
```


```{r}
monster_data_maxac |> 
  #create graph of HP vs AC for each monster
  ggplot(aes(x=AC, y=HP, color=Level, marker='x'))+ 
  #make scatter plot
  geom_point() + 
  #adjust color map for scale 
  scale_fill_brewer(palette = "RdYlBu")
```
Create a visual that includes the monster HP information with simulation data 

```{r}
#visualize the total_dmg output for each class split by ancestry 
#visualize the total_dmg output for each class split by ancestry 
class_select_sim |> ggplot(aes(x=target_ac, y=total_dmg)) +  
  geom_smooth(aes(group=ancestry, color=ancestry), method='lm', se=TRUE) +
  geom_point(data = monster_data_maxac, aes(x=AC, y=HP, shape=Level, size= Level, alpha=0.4)) + scale_shape_binned() +
    labs(title=glue('{class_choice} damage output by level and target ac'), 
       x='target ac', 
       y= 'total damage over 5 rounds')
       
```

Categorize monsters by whether they are BELOW class "moderate" curve or above 

```{r}

class_select_sim |> head()
```
Clean up monster spreadsheet
```{r}
monster_data_maxac <- monster_data_maxac |> select(-X)
monster_data_maxac |> head()
```

