R modeling to prepare code for R shiny app exploring monster encounter information with pathfinder 2e 
```{r}
library(tidyverse)
library(plotly)
```

```{r}
sim_results <- read.csv('../output_data/simualtion_results_ac10-30.csv')
sim_results|> head(3)
```
Identify rows or columns with blank or inaccurately copied filter columns and drop them (for now)
```{r}
# Replace all empty strings with NA
sim_results <- sim_results  |> mutate(across(c(ancestry, class_build), na_if, ""))

# sim_results |> drop_na(class_build) |> drop_na(ancestry)

sim_results |> distinct(class_build)
```
```{r}
all_class_options <- sim_results|> distinct(class_build) |> pull()
all_class_options

sim_results |> filter(class_build == all_class_options[1])
```


```{r}
# #make a visual where each class shows total damage 
# class_list <- sim_results |> c(distinct(class_build))
# 
# class_list |> pull(class_build)

### list all class options for chosing 
all_class_options <- sim_results|> distinct(class_build) |> pull()

class_choice = all_class_options[4]

#graph results from only one class ### figure out how to automate this process later 
class_select_sim <- sim_results |> filter(class_build == class_choice)

#visualize the total_dmg output for each class split by ancestry 
plot_test <- class_select_sim |> ggplot(aes(x=target_ac, y=total_dmg, color=ancestry)) +  
  geom_smooth(aes(group=ancestry), method='lm', se=TRUE) +
  labs(title=glue('{class_choice} damage output by level and target ac'), 
       x='target ac', 
       y= 'total damage over 5 round')

ggplotly(plot_test
)

```

troublehsoot ancestry plot color issue -> make it a factor variable?
```{r}
class_select_sim$ancestry <- factor(class_select_sim$ancestry)
class_select_sim
```

Open monster stat and add the monster HP distribution to the graph 
```{r}

monster_data<- read.csv('../output_data/bestiary_stats.csv')
monster_data|> head(3)

```

Visualize the distribution of total damage needed for specific hitpoint levels
```{r}
#filter so monster data only goes to the max AC level of simulation dataset 
max_target_ac <- sim_results |> distinct(target_ac) |> max()

monster_data_maxac <- monster_data |> filter(AC <= max_target_ac)

monster_data_maxac|> head()
```
```{r}
monster_data_maxac |> filter(AC == 12, HP == 20)

```


```{r}
monster_data_maxac |> 
  #create graph of HP vs AC for each monster
  ggplot(aes(x=AC, y=HP, color=Level, marker='x'))+ 
  #make scatter plot
  geom_point() + 
  #adjust color map for scale 
  scale_fill_brewer(palette = "RdYlBu")
```
Create a visual that includes the monster HP information with simulation data 

```{r}
#visualize the total_dmg output for each class split by ancestry 
#visualize the total_dmg output for each class split by ancestry 
class_select_sim |> ggplot(aes(x=target_ac, y=total_dmg)) +  
  geom_smooth(aes(group=ancestry, color=ancestry), method='lm', se=TRUE) +
  geom_point(data = monster_data_maxac, aes(x=AC, y=HP, shape=Level, size= Level, alpha=0.4)) + scale_shape_binned() +
  labs(title=glue('{class_choice} damage output by level and target ac'), 
       x='target ac', 
       y= 'total damage over 5 rounds')

```

Categorize monsters by whether they are BELOW class "moderate" curve or above 

```{r}

class_select_sim |> head()
```
Clean up monster spreadsheet
```{r}
monster_data_maxac <- monster_data_maxac |> select(-X)
monster_data_maxac |> head()
```


##############################
Combining the output damage of 4 selected party member combinations 
* below minimum -> trivial
* below 25th percentile -> easy
* 25th to 75th -> moderate
* 75th to max -> hard
* max and above -> impossible 

```{r}
### select party combinations 

anc_select_party = c('dwarf', 'ocr-str', 'halfling', 'elf')
class_select_party= c('barbarian', 'fighter', 'rogue','sorcerer')

## create a combination based on the location 

build_select_party = str_c(anc_select_party, class_select_party, sep = " ")
#print(build_select_party)

## Or create the list from a full build_name variable 

full_build_select_party =c("dwarf barbarian strength", "human-str fighter strength", "halfling rogue dexterity", "elf sorcerer charisma")



#####filter out these builds from  from the simulation data 

only_party_sim <- sim_results |> filter(build_name %in% full_build_select_party)

#Filter out the simulation dataset for each class individually 
class1_plot <- sim_results |>
  #Filter based on the class choice
  filter(build_name == full_build_select_party[1])
class2_plot <- sim_results |>
  #Filter based on the class choice
  filter(build_name== full_build_select_party[2])
class3_plot <- sim_results |>
  #Filter based on the class choice
  filter(build_name == full_build_select_party[3])
class4_plot <- sim_results |>
  #Filter based on the class choice
  filter(build_name == full_build_select_party[4])

#for each target_ac, add up all the damage from EACH encounter across the whole party 
## groupby by encounter number, than sum all the 'total damage'

party_combo_sim <- only_party_sim |> group_by(target_ac, encounter_num) |> summarise(total_party_damage= sum(total_dmg))
print(party_combo_sim |> head(3))

## Create summary dataframe based on summary statistics
total_dmg_dist <- party_combo_sim |>
  group_by(target_ac) |>
  summarise(min_val = min(total_party_damage),
            max_val = max(total_party_damage),
            q25= quantile(total_party_damage, 0.25),
            median= median(total_party_damage),
            q75= quantile(total_party_damage, 0.75))

### Create monster challenge rating based on where the HP falls on the total_damage distribution
#rename monster column for easier merging
monster_data_merge <- monster_data_maxac |>
  rename(target_ac = AC)

#merge the summary stats with monster info based on AC
mon_with_party_stat <- left_join(monster_data_merge,
                                 total_dmg_dist,
                                 by = "target_ac")

#create a challenge rating using case_when based on the summary statistics
mon_party_challenge <- mon_with_party_stat |>
  mutate(challenge_rate = case_when (HP<q25 ~ 'easy',
                                     HP >=q25 & HP <= q75 ~ 'moderate',
                                     HP > q75 & HP <= (2*max_val)~ 'hard',
                                     HP >= (2*max_val) ~'impossible')
  )

#convert challenge rating to a factor value
mon_party_challenge <- mon_party_challenge|>
  mutate (challenge_rate= as.factor(challenge_rate))

#create a color scale for the graphing
color_dict= c('easy'='green', 'moderate'='blue', 'hard'='red', 'impossible'='orange')
title <- glue("Monster challenge rating based on 4-party total damage output over 5 rounds")

## Graph where total_damage and target_ac is on y and x axis, and shows total distribtuion across all encounters

party_combo_sim |> ggplot(aes(x=target_ac, y=total_party_damage))  +
  geom_point(alpha = 1/10) +
  geom_smooth(method='lm', se=TRUE) + #add geom_ribbon() for predictive ranges
  #apply monster information to the plot -> ADD that the color is equal to the CHALLENGE RATING, and that the shape is equal to the level
  
  
   geom_point(data=mon_party_challenge,
                 aes(label= Name, 
                     x=target_ac,
                     y=HP,
                     shape= 13,
                     size= Level,
                     color= challenge_rate)) +
      scale_shape_binned() + 
  
  scale_color_manual(values = color_dict) +
  labs(title= title,
       x='target ac',
       y= 'target hit points') + 
  geom_smooth(data = class1_plot, aes(x=target_ac, y=total_dmg), 
              method='lm',
              se=TRUE) 

  #add in a lineplot for each class individually
  # #class 1
  # geom_smooth(data = class1_plot,
  #             method='lm',
  #             se=TRUE) +
  # # #class 2
  # geom_smooth(data = class2_plot,
  #             method='lm',
  #             se=TRUE) +
  # # #class 3
  # geom_smooth(data = class3_plot,
  #             method='lm',
  #             se=TRUE) +
  # # #class 4
  # geom_smooth(data = class4_plot,
  #             method='lm',
  #             se=TRUE) +



```

```{r}
class1_plot |> head()

mon_party_challenge |> head()
```

```{r}
party_combo_sim |> ggplot(aes(target_ac, total_party_damage)) + geom_point()
```


```{r}
total_dmg_dist <- party_combo_sim |> 
  group_by(target_ac) |> 
  summarize(min_val = min(total_party_damage), 
            max_val = max(total_party_damage), 
            q25= quantile(total_party_damage, 0.25), 
            median= median(total_party_damage), 
            q75= quantile(total_party_damage, 0.75))

#use a case_when function to classify the challenge rating based on where it falls on the total_dmg_dist

#rename monster column for easier merging 
monster_data_merge <- monster_data_maxac |> 
  rename(target_ac = AC)

#merge the summary stats with monster info based on AC
mon_with_party_stat <- left_join(monster_data_merge, 
                                 total_dmg_dist, 
                                 by = "target_ac")

#create a challenge rating using case_when based on the summary statistics 
mon_party_challenge <- mon_with_party_stat |> 
  mutate(challenge_rate = case_when (HP<q25 ~ 'easy', 
                                     HP >=q25 & HP <= q75 ~ 'moderate',
                                     HP > q75 & HP <= (2*max_val)~ 'hard', 
                                     HP >= (2*max_val) ~'impossible')
  )

#convert challenge rating to a factor value 
mon_party_challenge <- mon_party_challenge|> 
  mutate (challenge_rate= as.factor(challenge_rate))


```

```{r}


#use a case_when function to classify the challenge rating based on where it falls on the total_dmg_dist

# monster_data_maxac |> mutate(challenge_rate = case_when (AC ==total_dmg_dist$target_ac & HP <)
monster_data_merge <- monster_data_maxac |> rename(target_ac = AC)
monster_data_merge

#merge the summary stats with monster info based on AC
mon_with_party_stat <- left_join(monster_data_merge, total_dmg_dist, by = "target_ac")

#create a case_when based on the summary statistics 
mon_party_challenge <- mon_with_party_stat |> 
  mutate(challenge_rate = case_when (HP<q25 ~ 'easy', 
                                     HP >=q25 & HP <= q75 ~ 'moderate',
                                     HP > q75 & HP <= (2*max_val)~ 'hard', 
                                     HP >= (2*max_val) ~'impossible')
  )

#convert challenge rating to a factor value 
mon_party_challenge <- mon_party_challenge|> 
  mutate (challenge_rate= as.factor(challenge_rate))


```

```{r}
party_combo_sim |> ggplot( aes(x=target_ac, y=total_party_damage))  + 
  geom_point(alpha = 1/10) + 
  geom_smooth(method='lm', se=TRUE) +
  geom_quantile(quantiles=c(0.25, 0.75), color="gray") #add geom_ribbon() for predictive ranges 
#apply monster information to the plot -> ADD that the color is equal to the CHALLENGE RATING, and that the shape is equal to the level
```


```{r}
total_dmg_dist
```


```{r}
monster_challenge <- mon_party_challenge |> relocate(challenge_rate) |> rename("challenge rating"= challenge_rate) 


monster_challenge|> select(-c(min_val, max_val, q25, median, q75)) |> rename("AC" = target_ac)
                           

```

